---
title: "Volcano Plot"
author: "Jens Hooge"
date: 2015-07-23T21:13:14-05:00
categories: ["R"]
tags: ["R", "Visualization", "Library", "BioViz"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Volcano Plot

A volcano plot is a type of scatter-plot that is used to quickly identify changes in large data sets composed of replicate data. It plots significance versus fold-change on the y and x axes, respectively. These plots are increasingly common in omic experiments such as genomics, proteomics, and metabolomics where one often has a list of many thousands of replicate data points between two conditions and one wishes to quickly identify the most meaningful changes. A volcano plot combines a measure of statistical significance from a statistical test (e.g., a p value from an ANOVA model) with the magnitude of the change, enabling quick visual identification of those data-points (genes, etc.) that display large magnitude changes that are also statistically significant.

```{r volcano_plot, fig.cap='A fancy volcano plot', tidy=FALSE}
library(BioViz)
data(iPSC)
general.volcano_plot(iPSC, title = "Volcano Plot", groups = c("Untreated",
  "Treated"), top_labeled = 0, xlabel = expression(log[2]("Foldchange")),
  ylabel = expression(-log[10]("p-value")), xcutoff = c(-log2(4), log2(4)),
  ycutoff = -log10(0.05), log1p = FALSE)
```

# Including Plots

You can also embed plots. See Figure \@ref(fig:pie) for example:

```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```
