---
title: "Volcano Plot"
author: "Jens Hooge"
date: 2017-07-07
thumbnail: "images/volcano.png"
categories: ["general", "visualization"]
tags: ["Significance", "p-value", "Expression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Volcano Plot

A volcano plot is a type of scatter-plot that is used to quickly identify changes in large data sets composed of replicate data. It plots significance versus fold-change on the y and x axes, respectively. These plots are increasingly common in omic experiments such as genomics, proteomics, and metabolomics where one often has a list of many thousands of replicate data points between two conditions and one wishes to quickly identify the most meaningful changes. A volcano plot combines a measure of statistical significance from a statistical test (e.g., a p value from an ANOVA model) with the magnitude of the change, enabling quick visual identification of those data-points (genes, etc.) that display large magnitude changes that are also statistically significant.

## Simple
```{r simple, tidy=FALSE}
library(BioViz)
data(iPSC)
general.volcano_plot(iPSC)
```

## Volcano Plot with Group Labels
```{r, group_labels, tidy=FALSE}
general.volcano_plot(iPSC, groups=c("A", "B"))
```

## Volcano Plot annotated by top10 most significant samples
```{r annotated, tidy=FALSE}
general.volcano_plot(iPSC, top_labeled=10, xcutoff=c(-log2(2), log2(4)))
```

## Volcano Plot annotated by Top10 most Significant Samples and transformed y-axis by log(1+x)
```{r axis_transform, tidy=FALSE}
general.volcano_plot(iPSC, top_labeled=10, log1p=TRUE)
```
