---
title: Principle Components
author: Jens Hooge
date: '2017-07-12'
slug: principle-components
categories:
  - dimensionality_reduction
tags:
  - exploration
draft: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

## Component Plot
The components of the PCA dimensionality reduction will be displayed in a pairs plot. Each
of the components are plottet against each other. 

- The lower diagonal shows the data points projected to the respective two dimensional component space.
- The upper diagonal shows a contour plot of the data projected the respective two dimensional component space.
- The diagonal shows a density plot of the projection of variables to a principal component axis.

## Load data
```{r dataset, tidy=FALSE, warning=FALSE}
library(BioViz)
library(reshape2)
library(ggplot2)
data(fiveClass)
Y <- fiveClass$Class
X <- fiveClass[, 2:ncol(fiveClass)]
```

## Default Pairs Plot of the First 3 Principal Components
```{r errorbars, tidy=FALSE}
dimred.plot_pca(X)
```

## Pairs Plot of the First 3 Principal Components Labeled by Class Vector
```{r error_transform, tidy=FALSE}
dimred.plot_pca(X, Y)
```

## Varying the Number of Components
```{r facetted, tidy=FALSE}
dimred.plot_pca(X, Y, ncomp=5)
```
